---
title: "Dorm Fridge Data Analysis"
subtitle: "ME 331: Thermodynamics Spring 2026"
output: html_document
author: "Eva Bochner, Michelle Chen, Sierra Healy, Saina Srivastava"
date: "2026-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the data analysis for the data acquired from the Arduino for Spring 2026's ME 331 group Home Fridge.

## Baseline Data

First, the team imports important libraries. This includes tidyverse, a rudimentary package for statistical analysis, and tidymodels, a beginner-friendly package for statistical modeling.

```{r}
#| label: import-all-libraries
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(dplyr)
```

After importing these libraries, the team imported the data. The data is broken into two key chunks: the baseline data (*thermo_baseline.csv*) and the data collected over hours (*filename.csv).*

```{r}
#| label: import-all-data
#| message: FALSE
baseline <- read_csv("thermo_baseline.csv", show_col_types = FALSE)
baseline <- baseline |>
  select(where(~ !all(is.na(.))))
baseline
```

## 

The team then cleaned out the dataset so it only reflected the moments where the compressor in the fridge was using power, rather than the total data collection time, as many values for work rate (power) are 0.

```{r}
#| label: filtering-no-w

baseline_filt <- baseline |>
  filter(`W` != 0)
baseline_filt
```

Next, the team made two plots: minutes after start vs. W (work) and minutes after start vs. volts. **Figure 1** is the first plot.

```{r}
#| label: plot-work
ggplot(baseline_filt, aes(x = `Minutes after start`, y = W)) +
  geom_line()
```
