---
title: "Dorm Fridge Data Analysis"
subtitle: "ME 331: Thermodynamics Spring 2026"
output: html_document
author: "Eva Bochner, Michelle Chen, Sierra Healy, Saina Srivastava"
date: "2026-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the data analysis for the data acquired from the Arduino for Spring 2026's ME 331 group Home Fridge.

## Baseline Data

First, the team imports important libraries. This includes tidyverse, a rudimentary package for statistical analysis, and tidymodels, a beginner-friendly package for statistical modeling.

```{r}
#| label: import-all-libraries
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(dplyr)
```

After importing these libraries, the team imported the data. The data is broken into two key chunks: the baseline data (*thermo_baseline.csv*) and the data collected over hours (*trial1.csv, trial2.csv, trial3.csv).*

```{r}
#| label: import-all-data
#| message: FALSE
#| warning: FALSE
baseline <- read_csv("thermo_baseline.csv", show_col_types = FALSE)
baseline <- baseline |>
  select(where(~ !all(is.na(.)))) |>
  rename(time = `Minutes after start`)
trial1 <- read_csv("trial1.csv", col_types = cols_only(
  `Minutes after start` = col_double(), 
  `W` = col_double(),
  `temp in fridge` = col_double(),
  `temp in water` = col_double(),
  `temp in air` = col_double()), 
  show_col_types = FALSE, 
  name_repair = "unique_quiet") |>
  rename(time = `Minutes after start`,
         tfrid = `temp in fridge`,
         tw = `temp in water`,
         ta = `temp in air`,
         w_dot = `W`)
trial2 <- read_csv("trial2.csv", col_types = cols_only(
  `Minutes after start` = col_double(), 
  W = col_double(),
  `temp in fridge` = col_double(),
  `temp in water` = col_double(),
  `temp in air` = col_double()), 
  show_col_types = FALSE, 
  name_repair = "unique_quiet") |>
  rename(time = `Minutes after start`,
         tfrid = `temp in fridge`,
         tw = `temp in water`,
         ta = `temp in air`,
         w_dot = W)
trial3 <- read_csv("trial3.csv", col_types = cols_only(
  `Minutes after start` = col_double(), 
  W = col_double(),
  `temp in fridge` = col_double(),
  `temp in water` = col_double(),
  `temp in air` = col_double()), 
  show_col_types = FALSE, 
  name_repair = "unique_quiet") |>
  rename(time = `Minutes after start`,
         tfrid = `temp in fridge`,
         tw = `temp in water`,
         ta = `temp in air`,
         w_dot = W)

trial1
```

## Cleaning Process

The team then cleaned out the dataset so it only reflected the moments where the compressor in the fridge was using power, rather than the total data collection time, as many values for work rate (power) are 0.

```{r}
#| label: filtering-no-w

baseline_filt <- baseline |>
  filter(`W` != 0)
baseline_filt
```

## Plotting Baseline Data

Next, the team made two plots: minutes after start vs. W (work) and minutes after start vs. volts. **Figure 1** is the first plot.

```{r}
#| label: plot-work
ggplot(baseline_filt, aes(x = time, y = W)) +
  geom_line() +
  geom_point(color = "red") +
  labs(title = "Minutes after Start vs. the Power from the Wall Unit Over 90 Minutes",
       subtitle = "The power the compressor takes in in 5 minute bursts over time",
       x_label = "Time (min)", 
       y_lab = "Power (Watts)")
```

## Using Water Dish

After a clear idea, the team plotted the temperature of the air outside the fridge, water, and the air inside the fridge.

```{r}
#| label: plot-temp
#| warning: false
ggplot(trial1, aes(x = `time`)) +
 # geom_line(aes(y = `tfrid`)) +
  geom_line(aes(y = `tw`)) +
  #geom_line(aes(y = `ta`)) +
  labs(title = "Minutes after Start vs. the Temperature of the water",
       x = "Time (min)", 
       y = "Temperature (Celsius)")
```

Then, using the equation for heat transfer:

$$
\dot{Q} = \dot{m}c_v\Delta{T}
$$

By assuming that the heat transfer from the water is the same as the heat transferred into the refrigerator air, we can just observe the change in temperature of the water and then extrapolate those values for the fridge. The mass of the water is determined from the volume of the container, 250mL, and the density of water, which is 1 kg/m^3^.

So it would be:

$$
m = V \times\rho_w 
$$

Which in this case yields

$$m$$ = 0.25 kg. We use this to find the heat lost from the water due to refrigeration, which will then be divided by the time taken, to find the heat removal rate. Below is the heat removal rate and the heat removed.

```{r}
#| label: find-q-dot
q_water_1 <- trial1 |>
  mutate(q_w = 0.25 * (tw- lag(tw)) * 4.184) |>
  mutate(q_w_dot = q_w/(time * 60))
q_water_2 <- trial2 |>
  mutate(q_w = 0.25 * (tw- lag(tw)) * 4.184) |>
  mutate(q_w_dot = q_w/(time * 60))
q_water_3 <- trial3 |>
  mutate(q_w = 0.25 * (tw- lag(tw)) * 4.184)|>
  mutate(q_w_dot = q_w/(time * 60))
q_water_1
```

When this is graphed out:\

```{r}
#| label: graphed-q-dot
#| warning: false
ggplot() +
  geom_line(data = q_water_1, aes(x = time, y = q_w, color = "Trial 1")) +
  geom_line(data = q_water_2, aes(x = time, y = q_w, color = "Trial 2")) +
  geom_line(data = q_water_3, aes(x = time, y = q_w, color = "Trial 3")) +
  labs(title = "Minutes after Start vs. the Heat Removed From Water",
       x = "Time (min)", 
       y = "Heat Removal (kJ)") +
  theme_minimal()

ggplot() +
  geom_line(data = q_water_1, aes(x = time, y = q_w_dot, color = "Trial 1")) +
  geom_line(data = q_water_2, aes(x = time, y = q_w_dot, color = "Trial 2")) +
  geom_line(data = q_water_3, aes(x = time, y = q_w_dot, color = "Trial 3")) +
  labs(title = "Minutes after Start vs. the Heat Removal Rate of Water",
       x = "Time (min)", 
       y = "Heat Removal (kJ)") +
  theme_minimal()
```

We do the same for the air inside the fridge. To find the $$\dot{Q_a}$$ , we use the specific heat value of dry air, 1005 J/kg, assuming that there is only dry air in the fridge. We use the same formula as above to determine $$\dot{Q}$$. To determine the mass of the air, we used the same thing as above: the volume of the fridge multiplies by the density of dry air, which ended up being 2.2 feet^3^, which is 0.062 m^3^. With the dry air density being 1.293 kg/m^3^ at standard state, the mass becomes 0.0801 kg.

```{r}
#| label: finding-fridge-q-dot
q_frid1 <- trial1 |>
  mutate(q_f = 0.0801 * (tfrid - lag(tfrid)) * 1.005) |>
  mutate(q_f_dot = q_f/(time * 60))
q_frid2 <- trial2 |>
  mutate(q_f = 0.0801 * (tfrid - lag(tfrid)) * 1.005) |>
  mutate(q_f_dot = q_f/(time * 60))
q_frid3 <- trial3 |>
  mutate(q_f = 0.0801 * (tfrid - lag(tfrid)) * 1.005) |>
  mutate(q_f_dot = q_f/(time * 60))
```

Plotted this out:

```{r}
#| label: graphed-frid-q-dot
#| warning: false
ggplot() +
  geom_line(data = q_frid1, aes(x = time, y = q_f, color = "Trial 1")) +
  geom_line(data = q_frid2, aes(x = time, y = q_f, color = "Trial 2")) +
  geom_line(data = q_frid3, aes(x = time, y = q_f, color = "Trial 3")) +
  labs(title = "Minutes after Start vs. the Heat Removed from Air",
       x = "Time (min)", 
       y = "Heat Removed (kJ)") +
  theme_minimal()
```

The total work is found through integrating the work rate (power) by time.

```{r}
#| label: find-work
w1 <- trial1 |>
  mutate (w_dot = replace_na(w_dot, 0)) |>
  mutate(w = w_dot * (time - lag(time)) * 60 * 0.001) |> #makes this in kW
  select(time, w, w_dot)
w2 <- trial2 |>
  mutate (w_dot = replace_na(w_dot, 0)) |>
  mutate(w = w_dot * (time - lag(time)) * 60 * 0.001) |>
  select (time, w, w_dot)
w3 <- trial3 |>
  mutate (w_dot = replace_na(w_dot, 0)) |>
  mutate(w = w_dot * (time - lag(time)) * 60 * 0.001) |>
  select (time, w, w_dot)
w1
```

Plot work:

```{r}
#| label: graphed-work
#| warning: false
ggplot() +
  geom_line(data = w1, aes(x = time, y = w, color = "Trial 1")) +
  geom_line(data = w2, aes(x = time, y = w, color = "Trial 2")) +
  geom_line(data = w3, aes(x = time, y = w, color = "Trial 3")) +
  labs(title = "Minutes after Start vs. the Electric Work done",
       x = "Time (min)", 
       y = "Work (kJ)") +
  theme_minimal()
```

Putting all the data together:

```{r}
#| label: add-q-together
combined_q_1 <- q_water_1 |>
  full_join(q_frid1, by="time") |>
  full_join(w1, by="time") |>
  mutate(trial = "Trial 1") |>
  select(time, w_dot, w, q_f, q_w, q_f_dot, q_w_dot, trial)
combined_q_2 <- q_water_2 |>
  full_join(q_frid2, by="time") |>
  full_join(w2, by="time") |>
  mutate(trial = "Trial 2") |>
  select(time, w_dot, w, q_f, q_w, q_f_dot, q_w_dot, trial)
combined_q_3 <- q_water_3 |>
  full_join(q_frid3, by="time") |>
  full_join(w3, by="time") |>
  mutate(trial = "Trial 3") |>
  select(time, w_dot, w, q_f, q_w, q_f_dot, q_w_dot, trial)

all <- bind_rows(combined_q_1, combined_q_2, combined_q_3)
all[is.na(all)] <- 0
```

```{r}
all|>
  group_by(trial)|>
  summarize(q_total = sum(q_w_dot)) 
  ungroup(trial) |>
 summarize(q_lost_w = mean(q_w))
```

Using the first law of thermodynamics:

$$
Q - W~=~m\times\Delta{h}
$$

where $${m}$$ is 0 due to mass conservation, in an ideal scenario, the $$Q$$ equals the $$W$$. Thus, we combine $${Q_w}$$ with $$Q_a$$ to find $$Q_t$$, which is the total heat removed from the system due to the fridge. We do the same for the work to determine the total work and total heat per trial

```{r}
#| label: total-q-per-trial
total_q <- all |>
  group_by(trial) |>
  summarise(
    q_f_total = sum(q_f, na.rm = TRUE),
    q_w_total = sum(q_w, na.rm = TRUE),
    qt = sum(q_f, na.rm = TRUE) + sum(q_w, na.rm = TRUE),
    wt = sum(w, na.rm = TRUE)) 
total_q
```

After finding $$Q_t$$, we find the coefficient of performance (COP) using the formula:\

$$
COP~=~Q_{total}/W_{cycle}
$$

```{r}
#| label: cop
#| warning: FALSE
total_q |>
  summarise(cop = abs(qt)/wt, avg_cop = mean(cop))
```
