---
title: "Dorm Fridge Data Analysis"
subtitle: "ME 331: Thermodynamics Spring 2026"
output: html_document
author: "Eva Bochner, Michelle Chen, Sierra Healy, Saina Srivastava"
date: "2026-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the data analysis for the data acquired from the Arduino for Spring 2026's ME 331 group Home Fridge.

## Baseline Data

First, the team imports important libraries. This includes tidyverse, a rudimentary package for statistical analysis, and tidymodels, a beginner-friendly package for statistical modeling.

```{r}
#| label: import-all-libraries
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(dplyr)
```

After importing these libraries, the team imported the data. The data is broken into two key chunks: the baseline data (*thermo_baseline.csv*) and the data collected over hours (*filename.csv).*

```{r}
#| label: import-all-data
#| message: FALSE
baseline <- read_csv("thermo_baseline.csv", show_col_types = FALSE)
baseline <- baseline |>
  select(where(~ !all(is.na(.))))
trial1 <- read_csv("trial1.csv", col_types = cols_only(
  `Minutes after start` = col_double(), 
  `temp in fridge` = col_double(),
  `temp in water` = col_double(),
  `temp in air` = col_double()), 
  show_col_types = FALSE, 
  name_repair = "unique_quiet")

trial1
```

## Cleaning Process

The team then cleaned out the dataset so it only reflected the moments where the compressor in the fridge was using power, rather than the total data collection time, as many values for work rate (power) are 0.

```{r}
#| label: filtering-no-w

baseline_filt <- baseline |>
  filter(`W` != 0)
baseline_filt
```

## Plotting Baseline Data

Next, the team made two plots: minutes after start vs. W (work) and minutes after start vs. volts. **Figure 1** is the first plot.

```{r}
#| label: plot-work
ggplot(baseline_filt, aes(x = `Minutes after start`, y = W)) +
  geom_line() +
  geom_point(color = "red") +
  labs(title = "Minutes after Start vs. the Power from the Wall Unit Over 90 Minutes",
       subtitle = "The power the compressor takes in in 5 minute bursts over time",
       x_label = "Time (min)", 
       y_lab = "Power (Watts)")
```

## Using Water Dish

After a clear idea, the team placed a dish of water with the mass of *mass* to see the rate of heat transfer from water to the fridge. This variable will be called $$\dot{Q_w}$$

.Plotting this:

```{r}
#| label: plot-work
ggplot(trial1, aes(x = `Minutes after start`)) +
  geom_line(aes(y = `temp in fridge`)) +
  geom_line(aes(y = `temp in water`)) +
  geom_line(aes(y = `temp in air`)) +
  labs(title = "Minutes after Start vs. the Temperature",
       subtitle = "Measuring the temperature in air, water, and the fridge over time",
       x_label = "Time (min)", 
       y_lab = "Temperature (Celsius)")
```
