---
title: "Dorm Fridge Data Analysis"
subtitle: "ME 331: Thermodynamics Spring 2026"
output: html_document
author: "Eva Bochner, Michelle Chen, Sierra Healy, Saina Srivastava"
date: "2026-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the data analysis for the data acquired from the Arduino for Spring 2026's ME 331 group Home Fridge.

## Baseline Data

First, the team imports important libraries. This includes tidyverse, a rudimentary package for statistical analysis, and tidymodels, a beginner-friendly package for statistical modeling.

```{r}
#| label: import-all-libraries
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(dplyr)
```

After importing these libraries, the team imported the data. The data is broken into two key chunks: the baseline data (*thermo_baseline.csv*) and the data collected over hours (*filename.csv).*

```{r}
#| label: import-all-data
#| message: FALSE
baseline <- read_csv("thermo_baseline.csv", delimeter = ',')
baseline
baseline
```

## Including Plots

You can also embed plots, for example:

```{r}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
